---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# torpor

<!-- badges: start -->
<!-- badges: end -->

## Welcome to torpor

This is the repository of torpor an `R` (https://www.r-project.org/) package aiming at  an objective and standardized distinction between torpid and euthermic metabolic rates (MR) measured in steady-state conditions.
Note: This package is not aimed for general use without careful researcher’s attention.

## Installation

ToRpoR is using jags in the background and thus you should make sure to have [jags](http://mcmc-jags.sourceforge.net) installed 
on your machine. Once jags is properly installed and working you can installed torpor in R by typing:  

```{r eval=FALSE, include=TRUE}
library(remotes)
remotes::install_github("vullioud/torpor", build_opts =  c("--build_vignettes"),force=TRUE)
```

The package will be submitted to cran in the forseeable future. 

## Where can I find more about toRpoR?
You can learn more about the goal and the theory behind the package in the companion article (Fassel et al., in prep). 
