---
title: "An example"
author: "Colin Vullioud"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## How to use toRpoR ? An example

This vignette presents a suggested workflow of the `toRpoR` package.


### The data 
In order to fit the model, we need data. The input for the main function `fit_torpor()` consists of 
a set of measure of Metabolic rate $(M)$ and associated ambient temperature $(T_a)$.
These value can be vectors of the same length or columns of a data-frame. In this 
example, we will use the data taken from the measurment of Geiser (1987) that comes 
with the package. The data are accessible with the following command. 

```{r, data}
library(toRpoR)
data(test_data2) 
str(test_data2)
``` 
For this examples we thus get 103 observations of Vo2 of the species *Cercartetus lepidus*
and the ambiant temperature at which the recording have been made. 

$T_{lc}$ value and $BMR$ are also needed in order to fit the model. In the example
data-sets these values are found in the doccumentation of the data. We can find this by calling
the help function on the data set. In this example by typing: `?test_data2`. 
Once we get this information we are good to fit the model using the function
`fit_torpor()`. `fit_torpor()` represents the core of the model and should be the first
step in any analysis using the `toRpoR` package. Researchers not familiar with 
R and are only interested in a visual representation of their data can skip this 
step and use the `plot_torpor()` function directly but this is not recommended as 
problems might remain unnoticed. 

The model is fitted with the following call: 

## fit model 
```{r, fit model}
model <- fit_torpor(MR = test_data2[, "VO2ms"], Ta = test_data2[, "Ta"],
                    BMR = 1.49, TLC = 28.8,
                    fitting_options = list(ni = 5e+05,
                                           nt = 10,
                                           nb = 3e+05,
                                           nc = 2))
head(model)

# fake data -> warning check  overlap check 
# model2 <- fit_torpor(MR = test_data[, 2], Ta = test_data[,1])
# model2
```

The output of the `fit_torpor()` function is a list of class `jagsUI`. 
Researchers familiar with jagsUI are welcome to build their analysis from that point. 
However, the rest of the package is designed to help researcher perform common 
task and ease the process of making sens of the model. 


## Summary

The first step and the best way to have a quick view on the model is to call the
function `get_summary()` on the fitted model. This will return a list with the 
parameter estimates, their credible interval, the $\hat{R}$ value.... 

```{r, get_summary}
summary <- get_summary(model)

```

the parameter estimates are accessible by: 

```{r, parameter estimates}
summary$parameter_estimates
```

If the $\hat{R}$ values are distant from 1 we recommand to refit the model and 
increase the number of iteration and the burn-in. This is done by setting the 
parameter `fitting_options = list(ni = "something bigger", nb = "something bigger")`


and the overlap by: 

```{r, overlap}

summary$overlap
``` 


We recommand the user to first check the convergence. User can use the build in 
function from `jagsUI` to check convergence. We also implemented a function to 
check the overlap between the posterior and the prior distribution. We recommand
to check this before analysing the result further. 



## obtain classifcation and/or summary

## check convergence 
R hat mod_output -> summary > 1.1


plot possible avec jagsUI::traceplot(mod, "parameter")

si problem _> augmenter increase ni, nb. 

summary donne overlap aussi 

## check parameters identifiability (sugested by the authors)  
To continue with our analysis of *Cercartetus lepidus* we will check for overlap: 

```{r, check overlap}
# overlap <- check_overlap(mod = model)
# overlap
``` 

The function `check_overlap()` only check overlap for the $T_{lc}$, $T_t$, and 
$\beta_t$. It will return a warning if the overlap is greater between 0.3.
## Evaluate output 

prediction // classification // plot

## plot the data and prediction 


##



