---
title: "An example"
author: "Colin Vullioud"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## How to use torpor ? An example

This vignette presents a suggested workflow of the `torpor` package. It is aimed 
in priority at users not familiar with `rjags` yet interested in understanding 
the thermo-regulatory pattern of their favorit species. 

This example is based on one of the example datasets. We go through a complete 
example on how to use the functions of the package and in which 
order. At each step we also presents the options for the users and the possible 
solution to common problems. Note however, that this examples his not an exhaustive presentation
of the package. 


### The data 
The first step in a analysis is to get data. In order to fit a model using the
function `fit_torpor()` we need a set of measure of Metabolic rate $(M)$ and 
the ambient temperature $(T_a)$ at which the measurment have been made. 
These value can be vectors of the same length or two columns of a data-frame. In this 
example, we will use the data taken from the measurment of Geiser (1987) of the 
species *Cercartetus lepidus*.

The data are accessible with the following command. 

```{r, data}
library(torpor)
data(test_data2) 
str(test_data2)
``` 
We get 103 observations of Vo2 of the species *Cercartetus lepidus*

The lower critical temperature (i.e., lower limit of the thermoneutral zone, $T_{lc}$) 
and the basal metabolic rate ($BMR$) are also needed in order to fit the model with 
the `fit_torpor()` function. For the package's datasets, these values are found in 
the doccumentation of the data. We can find this by calling the help function on
the data set. For this example we access them by typing: `?test_data2`. 

`fit_torpor()` represents the core function of the model and should be the first
step in any analysis using the `torpor` package. Researchers not familiar with 
R and only interested in a visual representation of their data can skip this 
step and use the `plot_torpor()` function directly. This option is however not 
recommended as problems might remain unnoticed. 

The model is fitted with the following call: 

## fit model 
```{r, fit model}
model <- fit_torpor(MR = test_data2[, "VO2ms"], Ta = test_data2[, "Ta"],
                    BMR = 1.49, TLC = 28.8,
                    fitting_options = list(ni = 5e+03,
                                           nt = 10,
                                           nb = 3e+03,
                                           nc = 2))
```

Note that for the purpose of this vignette we only run 5000 iterations with 3000 
burned in. The results are thus not trusthworthy.

The output of the `fit_torpor()` function is a list of class `jagsUI`. This can be
intimidating for the user not familiar with this package. Therefore, the rest of the package offers
several functions to make sens of the model, check its consistency and represent it graphically. 
Researchers familiar with jagsUI can build their analysis from that point. 


## Summary

Once the model if fitted, it is recommanded to have an overlook at the results. 
This can be achieved via the function `get_summary()`. This will return a list with the essentials:
A tibble with the mean and median of parameter estimates, the 95 credible interval and the $\hat{R}$ value.
It also reports the parameters identification. 

```{r, get_summary}
summary <- get_summary(model)
```

the parameter estimates are accessible with: 

```{r, parameter estimates}
summary$parameter_estimates
```

### Check convergence
If the $\hat{R}$ values, given in the summary tibble are distant **to be precised** from 1 we recommand to refit the model and 
increase the number of iteration and burn-in respectively. This can be done by setting the 
parameter `fitting_options = list(ni = , nb = )`

It is also possible to look graphicaly at the convergence using the `jagsUI` 
function `traceplot()`. Let's have a look at the convergence for the parameters 
`betat` and `betac`. 

```{r, traceplot betat}
jagsUI::traceplot(model, c("betat", "betac"))
```

## check parameters identifiability 
In addition to check of the convergence it is advised, in order to trust the 
results of the model, to test the identifiability of the parameters. This is done
by comparing the prior distribution to the posterior. If the overlap is greater 
than 30% the parameter is considered to be not identified and a warning will be 
sent to the user. The parameter identifiability is done for the $T_{lc}$, $T_t$, and 
$\beta_t$. 

The ovelap is also given by the `get_summary()` function. Let's continue with 
our analysis of *Cercartetus lepidus*: 

```{r message=TRUE, warning=TRUE, paged.print=FALSE}
summary$overlap
``` 

## Evaluate output 

Once the basic check are done, we can go on to evalutation of the results. There are 
two main functions to deal with predictions. The first gives the classification for
the raw data based on the predictions of the model (`get_classification()`), the other 
give the predicted $MR$ in torpor and euthermy for a given $T_a$ (`get_prediction()`). 
Finaly we can plot the result using the function `plot_torpor()`


### Classification 

To look at the classification and the prediction for our data we use the function `get_classification()`
which will classify the measured points either in torpor or eutermy and returns a 
tibble with the classification (predicted_stage), the predicted $M_r$, the measured $M_r$
and the measured $T_a$. 

```{r, get_classification}
classification <- get_classification(model)

head(classification)
```



### Prediction 
The `get_prediction()` function is slightly different as it takes a vector of $T_a$ 
as input and return the predicted $M_r$ both in torpor and in euthermy and the 95 credible interval. 
For example let's see what are the predicted $M_r$ for a $T_a$ of 22°C. 

```{r, get_prediction}
prediction <- get_prediction(model, 22)

head(prediction)
```

### plot the data and prediction

finally a built-in function allow the user to plot the results. A few parameters 
can be modulated by the user directly (labels with `xlab` and `ylab`, colors with `col_torp` and `col_eut`). 
It is also possible to save the plot using the arguments `pdf = TRUE`. Finaly users 
can choose if the prefer a base-R plot or a ggplot via the `plot_type` argument. This function 
required arguments are provided. 

```{r, plot}
plot <- plot_torpor(mod = model, plot_type = "ggplot", ylab = "MR")
plot 

plot2 <- plot_torpor(MR = test_data2[, "VO2ms"], Ta = test_data2[, "Ta"],
                    BMR = 1.49, TLC = 28.8,
                    fitting_options = list(ni = 5e+03,
                                           nt = 10,
                                           nb = 3e+03,
                                           nc = 2), 
                    plot_type = "ggplot", 
                    ylab = "MR")
plot2

plot3 <- plot_torpor(mod = model, plot_type = "base", ylab = "MR")
plot3


```


The plot can be directly saved as a pdf using the argument `pdf = TRUE` or custom
further using either ggplot or graphic. 


```{r, plot custom}
plot + ggplot2::ggtitle("Cercartetus lepidus") +
  ggplot2::ylab("Test") +
  ggplot2::theme_minimal() +
  ggplot2::theme(legend.position="bottom")
```


