---
title: "An example"
author: "Colin Vullioud"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## An example 

This vignette presents a suggested workflow of the `toRpoR` package.

### The data 
The model needs data to fit with the function `fit_torpor()` consists of 
a set of measure of Metabolic rate $(M)$ and associated ambient temperature $(T_a)$.
These value can be vectors of the same length or columns of a data-frame. In this 
example, we will use the data taken from the measurment of Geiser (1987) that comes 
with the package. The data are accessible with the following command. 

```{r, data}
library(toRpoR)
data(test_data2) 
str(test_data2)
dim(test_data2)
``` 
For this examples we thus get 103 observations of Vo2 of the bat species *Cercartetus lepidus*
and the ambiant temperature at which the recording have been done. 

## get data 
$T_{lc}$ value and $BMR$ are also needed in order to fit the model. In the example
data-set ths is found in the doccumentation of the data. You can find this by calling
the help function on the data set. In this example by typing: `?test_data2`. 
Once you get this information you are good to fit the model using the function
`fit_torpor()`. `fit_torpor()` represents the core of the model and should be the first
step in any analysis using the `toRpoR` package. For researcher not familiar with 
R and are only interested in a visual representation of their data can skip this 
step and use the `plot_torpor()` function directly. 

Now that we get the information we need, we can fit the model with the following 
call: 

## fit model 
```{r, fit model}
# model <- fit_torpor(MR = test_data2[, "VO2ms"], Ta = test_data2[, "Ta"],
#                     BMR = 1.49, TLC = 28.8,
#                     fitting_options = list(ni = 5e+05,
#                                            nt = 10,
#                                            nb = 3e+05,
#                                            nc = 2))
# 
# model2 <- fit_torpor(MR = test_data[, 2], Ta = test_data[,1])

```

The output of the `fit_torpor()` function is a list of class jagsUI. It can be 
usefull for the researchers familiar with jagsUI and they are welcome to build their 
analysis from that point. However, the rest of the package is designed to help researcher 
to perform common task and ease the process of making sens of the model. 

We recommand the user to first check the convergence. User can use the build in 
function from `jagsUI` to check convergence. We also implemented a function to 
check the overlap between the posterior and the prior distribution. We recommand
to check this before analysing the result further. 


## check overlap 
To continue with our analysis of *Cercartetus lepidus* we will check for overlap: 

```{r, check overlap}
# overlap <- check_overlap(mod = model)
# overlap
``` 

The function `check_overlap()` only check overlap for the $T_{lc}$, $T_t$, and 
$\beta_t$. It will return a warning if the overlap is greater between 0.3. 
 
##



